### 2011-02-10 paper/hyperNeatMethod.tex
HN edits, other

HyperNEAT is capable of exploiting the geometry of a problem because
the link values between nodes in the final ANN phenotype are a
function of the geometric positions of those nodes~\cite{stanley2009hypercube, clune2009sensitivity,
  clune2011performance}. In the case of quadruped locomotion, this
property has previously been shown to help Hyper-NEAT produce gaits with front-back, left-right,
and four-way symmetries~\cite{clune2009evolving,
  clune2011performance}.

@@

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current angles of each of the 9 joints of the robot
(described below) and a sine and cosine wave (to facilitate the
production of periodic behaviors). The sine and cosine waves were the
following function of time ($t$) in milli-seconds:
$\sin(t/12)\pi$. This function produces numbers in the range [$-\pi,
  \pi$], which were then JASON FILL ME IN TO DESCRIBE HOW JOINT
COMMANDS WERE SPECIFIED.

\editbox{are we mentioning that we gave HyperNEAT the center joint
  here or earlier?}

\edit{jason: edit this section to reflect the post-processing...}

>>>>

HyperNEAT is capable of exploiting the geometry of a problem because
the link values between nodes in the final ANN phenotype are a
function of the geometric positions of those nodes~\cite{stanley2009hypercube, clune2009sensitivity,
  clune2011performance}. In the case of quadruped locomotion, this
property has previously been shown to help HyperNEAT produce gaits with front-back, left-right,
and four-way symmetries~\cite{clune2009evolving,
  clune2011performance}.

@@

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current angles of each of the 9 joints of the robot
(described below \edit{Jeff: is it?}) and a sine and cosine wave (to facilitate the
production of periodic behaviors). The sine and cosine waves were the
following function of time ($t$) in milli-seconds:
$\sin(t/12)\pi$ \editbox{Jeff: this can't be right, because it would imply a period of only 12*2*pi=75 ms. Perhaps this was as a function of time in units of timesteps?}. This function produces numbers in the range [$-\pi,
  \pi$] \edit{Jeff: this is actually [-1,1], right?}, which were then JASON FILL ME IN TO DESCRIBE HOW JOINT
COMMANDS WERE SPECIFIED.

\editbox{Jeff: proofread this paragraph to make sure I didn't lie (too
  much)}

The outputs of the substrate at each time step were nine numbers in
the range $[-1,1]$, which were scaled according to the allowable
ranges for each of the nine motors and then used as commanded
positions.  Occasionally HyperNEAT would produce networks that
exhibited rapid oscillatory behavior, switching between nearly -1 and
nearly 1 from one time step to the next.  This resulted in motor
commands to alternate extremes every 25ms (using a command rate of
40Hz), which tended to damage and overheat the motors.  To ameliorate
this problem, we simply requested four times as many commanded
positions from the ANN and averaged over four commands at a time to
obtain the actual $g(t)$ to be used as a gait.  This proved to work
well and did not restrict the expressiveness of HyperNEAT.

%\editbox{are we mentioning that we gave HyperNEAT the center joint
%  here or earlier?}

### 2011-02-10 paper/learningMethods.tex
HN edits, other

We evaluated a total of seven different
learning algorithms for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial of each run.  This could be done by selecting a random starting vector within the
allowable domain.  However, to most directly compare learning methods
by their ability to improve on the initially chosen $\vec{\theta}$, we
decided to evaluate the different methods by starting them at the same three
initial $\vec{\theta}$ vectors in the three runs.

@@

The seven learning algorithms for the parametrized motion models are as follows:

@@

Each of the seven learning methods was run on 3 different initial
parameter vectors ($\vec{\theta}_A$,
$\vec{\theta}_B$, and $\vec{\theta}_C$) in order to fairly compare the algorithms. Three runs were performed per learning method. Runs continued until the performance plateaued, which we defined as when there was no improvement during the last third of
a run.

>>>>

We evaluated a total of six different
learning algorithms for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial of each run.  This could be done by selecting a random starting vector within the
allowable domain.  However, to most directly compare learning methods
by their ability to improve on the initially chosen $\vec{\theta}$, we
decided to evaluate the different methods by starting them at the same three
initial $\vec{\theta}$ vectors in the three runs.

@@

The six learning algorithms for the parametrized motion models are as follows:

@@

Each of the six learning methods was run on 3 different initial
parameter vectors ($\vec{\theta}_A$,
$\vec{\theta}_B$, and $\vec{\theta}_C$) in order to fairly compare the algorithms. Three runs were performed per learning method. Runs continued until the performance plateaued, which we defined as when there was no improvement during the last third of
a run.
